\documentclass[twocolumn,12pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc} 
\usepackage[english]{babel} 

\pagestyle{empty} 


\usepackage{amsmath, amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem*{proof*}{Proof}
\newtheorem{definition}{Definition}

\newtheorem{exercise}{Exercise}
\newtheorem{question}{Question}

\newtheorem{example}{Example}

\newtheorem{remark}{Remark}

\usepackage[a4paper,left=.8cm, right=.8cm,top=1.5cm,bottom=1.5cm]{geometry}
\setlength{\columnsep}{1.2cm}

\usepackage{cancel}
\usepackage{bm}
\usepackage{amssymb,amsfonts}
\usepackage{mathrsfs}
\usepackage{color}
%\usepackage{hyperref}
\usepackage{dsfont}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{marginnote}
\newcommand{\Ptr}{\mathcal P^{\rm tr}}
\newcommand{\tr}{{\rm tr}}
\newcommand{\N}{\mathbb N}
\newcommand{\calN}{\mathcal N}
\newcommand{\bP}{\bold P}
\newcommand{\calK}{\mathcal K}
\newcommand{\calF}{\mathcal F}
\newcommand{\calH}{\mathcal H}
\newcommand{\calP}{\mathcal P}
\newcommand{\calC}{\mathcal C}
\newcommand\red[1]{\textcolor{red} {#1} }
\newcommand{\R}{\mathbb R}
\newcommand{\bX}{\bar X}
\newcommand{\Ktr}{\calK^{\rm tr}}
\title{ \bfseries \Huge {Handout 2: Random walks }}    
\vspace{-4cm}        

\date{Due date : February $13^{th}$}       
\vspace{-4cm}        
\newcounter{num}  % Create a new counter for paragraphs
\begin{document}
	\maketitle
	\setcounter{num}{1}  % Start the paragraph counter at 1
	
	\thispagestyle{empty} 
	\paragraph{Brief recap}
	\begin{enumerate}
		\item A stochastic process is a sequence of random variables indexed by time.
		Alternatively, a stochastic process is a probability distribution over a set of paths.
	\item 	Let \( Y_1, Y_2, \dots \) be i.i.d. rvs s.t. \( Y_i = \pm 1 \) with equal probability. Let \( X_0 = 0 \) and
	\begin{equation}\label{eq:rand_walk}
		X_k = Y_1 + Y_2 + \cdots + Y_k,
	\end{equation}
	for all \( k \geq 1 \). This gives a probability distribution over the sequences $( X_0, X_1, X_2, \dots)$, and thus defines a discrete-time stochastic process. This process is known as the 1D simple (or symmetric) random walk.
	By the CLT, for large enough $n$, the distribution of \( \frac{1}{\sqrt{n}} X_n \) converges to $\calN(0,1)$.
	
		\end{enumerate}


\paragraph{Exercise \thenum.}{\bfseries (The results in this exercise are a continuation of the lecture, they are part of the mandatory learning material).}
Consider the stochastic process defined in~\eqref{eq:rand_walk}.
Prove the following statements:
\begin{enumerate}
	\item \( \mathbb{E}[X_k] = 0 \) for all \( k\geq 0 \).
	\item (Independent increment) For all \( 0 = k_0 \leq k_1 \leq \dots \leq k_r \), the random variables \( X_{k_{i+1}} - X_{k_i} \) for \( 0 \leq i \leq r - 1 \) are mutually independent.
	\item (Stationary) For all \( h \geq 1 \) and \( k \geq 0 \), the distribution of \( X_{k+h} - X_k \) is the same as the distribution of \( X_h \).
\end{enumerate}



\stepcounter{num}
\paragraph{Exercise \thenum.}

A CC student leaves Riad 9 after class.
The student is walking along the Riads' walkable alley and is torn between making a trip to Palm's to chat with friends and going to the learning center to do homework. 
The student is $n$ steps from Palms' and $T-n$ steps from the learning center where $T \geq 3$ is the distance between Palms' and the
learning center, both located on opposite ends of the Riads' walkable alley.
Suddenly, the student remembers MCS123 course and decides to play a game for probabilistic intuition to sink in.
The student flips a coin at each step to decide which way to go. 
With probability $p$, the student takes 2 steps toward the restaurant and with probability $1-p$, they take 1 step toward the learning center. 
If the student is one or two steps away from Palms' and takes two steps toward it, then the student reaches Palms'. 
If they are one step away from the learning center and take a step in its direction, then they reach the learning center. 
Let $X_n$ be the probability that the student reaches the
restaurant before the learning center. 
Assume that all coin tosses are mutually independent.
Find a recurrence for $X_n$. (You do not need to solve the recurrence)

%X_n = p X_{n+2} + (1-p)X_{n-1}

\iffalse
\stepcounter{num} 
\paragraph{Exercise \thenum.}
For two positive integers \( A \) and \( B \), what is the probability that a symmetric 1D random walk reaches \( A \) before it reaches \( -B \)? Let \( \tau \) be the first time at which the random walk reaches either \( A \) or \( -B \). Then \( X_\tau = A \) or \( -B \).
\fi


\stepcounter{num} 
\paragraph{Exercise \thenum.}
An ant was hunting for food in a house.
While the ant was walking on a rule in a drawer, someone sprayed insecticide, and poor ant got dizzy. 
It moves $n$ steps on the rule. 
The ant starts at 0, and at each step it moves 1 mm to the right or to the left, with equal probabilities. 
Assume all steps are independent. 
Let $Z$ be the ant's position after $n$ steps. 
Find the PMF of $Z$. 
%If k is an integer between −n and n (inclusive) such that n+k is an even number, we get
%$$ P(Z = k) = \binom{n}{\frac{n+k}{2}}\left(\frac{1}{2}\right)^n$$
%otherwise $ P(Z = k) = 0$


\stepcounter{num} 
\paragraph{Exercise \thenum.}
(Gambler's ruin). Two gamblers, A and B, make a sequence of
1\$ bets. In each bet, gambler A has probability p of winning, and gambler B has probability $q = 1 - p$ of winning. Gambler A starts with i dollars and gambler B starts with $N - i$ dollars; the total wealth between the two remains constant since every time A loses a dollar, the dollar goes to B, and vice versa.
We can visualize this game as a random walk on the integers between 0 and N , where p is the probability of going to the right in a given step: imagine a person who starts at position i and, at each time step, moves one step to the right with probability $p$ and one step to the left with probability $q = 1 - p$. 
The game ends when either A or B is ruined, i.e., when the random walk reaches 0 or $N$. 
What is the probability that A wins the game (walking away with all the money)?

\iffalse
\stepcounter{num} 
\paragraph{Exercise \thenum.}
An immortal drunk man wanders around randomly on the integers. 
He starts at the origin, and at each step he moves 1 unit to the right or 1 unit to the left, with equal probabilities, independently of all
his previous steps. 
Let $b$ be a googolplex (this is $10g$, where $g = 10^{100}$ is a googol).
\begin{enumerate}
	\item Find a simple expression for the probability that the immortal drunk visits $b$ before returning to the origin for the first time.
	\item Find the expected number of times that the immortal drunk visits $b$ before returning to the origin for the first time.
\end{enumerate}
\fi

\paragraph{Acknowledgments: } J. Blitzstein, J. Huang, C. Lee, T. Leighton, M. Van Dijk.
\end{document}